AOD Collocated with MAIAC: New Variables

### 2019-07-30 Code/03.2_AOD Variation.Rmd
check within day variation

---
title: "AODV3 Level2 AOD2.0 from Aeronet: Day Variation"
author: "Yang Liu"
date: "7/24/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 8, fig.height = 4)
```

```{r}
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
source(here::here("Code", "funcs_for_plots.R"))
# compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit#

@@

@@

@output dataset has dailyN: observations with in a day (at least 20)
@d_mean and sd_rmse are daily mean and sd

@@

# year > 2000
aod_conus <- aod20_data_pred470[year(date)>=2000][site%in%aeronet_conus_dt$Site_Name][!is.na(AOD_470nm)]
# aod_conus <- aod_conus[year(date)>2000]
aod_conus[,date:=as.Date(date, "%Y-%m-%d")]
aod_conus[,dailyN:=.N, by = .(site, date)]
aod_conus <- aod_conus[dailyN>20,]

@@

# summary1: aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[,.(lat, lon, dailyN, season, mean_aod = mean(AOD_470nm), 
                                           sd_rmse = sd(AOD_470nm)), by = .(site, date)])

@@

## daily SD AOD generally follows daily mean AOD, more outliers in the summer

Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date.  
Plot by time gives this figure:

```{r}
# show daily trend: AOD vs date, all data
ggplot(data = aod1, aes(x = date, y = sd_rmse, color = site)) +
  geom_point(size = 1) + 
  theme_classic() + theme(legend.position = "none") + 
  labs(y = "Daily SD(AOD) by station", x = "Date") + 
  ggtitle("Daily variation by station across time", subtitle = "(# Daily obs >20)")

@@

Trend within day

```{r, fig.width = 12, fig.height = 10}
# limit to long stations
site1 <- aod_conus[,.N, by = site][N>3330, site]

d1 <- aod_conus[site%in%site1,.(lat, lon, hour, dailyN, season, AOD_470nm, 
                                MeanAOD = mean(AOD_470nm),
                                sd_rmse = sd(AOD_470nm)), by = .(site, date)]
d1[, stdAOD:=AOD_470nm - MeanAOD]
d1[, hour := as.numeric(hour)]
d1[hour%in%c(0,1,2), hour :=hour+24]
d1[, totalN := sum(dailyN), by = site]
setorder(d1, -totalN)
# data ready 
site2 <- aod_conus[,.N, by = site][N>8E4, site]
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = stdAOD, color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam") + 
  ggtitle("Daily variation (AOD - daily Mean AOD) by hour")

# sd(AOD) by hour, mean follows same trend
d1_hour <- d1[site%in%site2,.(sd_hour = sd(AOD_470nm), mean_hour = mean(AOD_470nm)), by = hour]
setkey(d1_hour, hour)
plot(x = d1_hour$hour, y = d1_hour$sd_hour, type = "o", ylab = "sd(AOD)")

#
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour")

#
ggplot(dilute.data(d1[site%in%site2,],10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour")  + 
   facet_wrap(~season, ncol = 2)

# all sites

ggplot(dilute.data(d1,10), aes(x = hour, y = log(AOD_470nm), color = site)) + 
         geom_point(size = 0.5, alpha = 0.5) + theme_classic() + 
         geom_smooth(method = "gam",se = F) + 
  ggtitle("Daily variation log(AOD) by hour") + 
  theme(legend.position = "none") + 
  facet_wrap(~season, ncol = 2)
```

## SD(AOD) by station

Then we calculated the median daily SD (very similar to mean) across all the days for each station.  
*N* adds up the number of all the hour-station observations _N = sum(dailyN)_  
mediansd = median(sd_rmse), meansd = mean(sd_rmse)

```{r}
aod2 <- unique(aod1[,.(mediansd = median(sd_rmse), N = sum(dailyN), lon, lat),
             by = .(site)])

setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2[N>5000 & mediansd > 0.02,])

N_label = 5E4
ggplot(data = aod2[N>3330,], aes(x = N, y = mediansd))+
    geom_point(size = 1, alpha = 1) + 
    theme_classic() + 
    ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = N, y = mediansd, label = site)) + 
    ggtitle("SD vs. Total Obs by Station", subtitle = ("Half the stations *")) + 
    labs(x = "Number of Observations (2000 - 2019)", y = "SD(AOD)")

@@

* I have removed half of the stations with low total observations, that removed only `r  round(aod2[N<3330,sum(N)]/aod2[,sum(N)])` % of the AERONET AOD measurements (hour-station observation).

Generally speaking, smaller sites have higher variation, but still there are also some long-standing sites that have higher variation than others.

```{r, fig.width = 8, fig.height = 5}

ggplot() +
        geom_polygon(data = map_data("state"), 
                     aes(x = long, y = lat, group = group), 
                     fill = 'NA', color = "grey") + 
        geom_point(data = aod2[N>3330,],
                   aes(x=lon, y=lat, color = mediansd, size = N),
                   alpha = 0.8) + 
        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
        scale_size_area(max_size = 12) + 
        scale_colour_viridis_c() + 
        labs(x = "Longitude", y = "Latitude", color= "RMSE", 
             size = "Number of\nObservations") +
        ggrepel::geom_label_repel(data = aod2[N>N_label&mediansd>0.015], aes(x = lon, y = lat, label = site)) + 
        theme_minimal() +
        ggtitle("SD by AERONET stations") +
        theme(legend.position ='right')

@@

## Scatterplot and map By season

```{r, fig.width = 12, fig.height = 10}
aod1[, month := factor(months(date, abbr = T),
                            levels = month.abb)]
  aod1[month%in%c("Dec","Jan","Feb"), season:="Winter"]
  aod1[month%in%c("Mar","Apr","May"), season:="Spring"]
  aod1[month%in%c("Jun","Jul","Aug"), season:="Summer"]
  aod1[month%in%c("Sep","Oct","Nov"), season:="Fall"]
  aod1[, season:= factor(season, levels = c("Spring","Summer", "Fall", "Winter"))]

aod3 <- unique(aod1[,.(mediansd = median(sd_rmse), meansd = mean(sd_rmse), N = sum(dailyN), lon, lat),
             by = .(site, season)])

ggplot(data = aod3[N>3330,], aes(x = N, y = mediansd))+
    geom_point(size = 1, alpha = 1) + 
    theme_classic() + 
    facet_wrap(~ season, ncol = 2) + 
    ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = N, y = mediansd, label = site)) + 
    ggtitle("SD vs. Total Obs by Station", 
            subtitle = ("Showing long-lasting stations, half stations, 97% of the observations")) + 
    labs(x = "Number of Observations (2000 - 2019)", y = "SD by station")

ggplot() +
        geom_polygon(data = map_data("state"), 
                     aes(x = long, y = lat, group = group), 
                     fill = 'NA', color = "grey") + 
        geom_point(data = aod3[N>3330,],
                   aes(x=lon, y=lat, color = mediansd, size = N),
                   alpha = 0.8) + 
        # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
        scale_size_area(max_size = 8) + 
        scale_colour_viridis_c() + 
        facet_wrap(~ season, ncol = 2) + 
        labs(x = "Longitude", y = "Latitude", color= "SD(AOD)", 
             size = "Number of\nObservations") +
        ggrepel::geom_label_repel(data = aod3[N>2E4&mediansd>0.02], aes(x = lon, y = lat, label = site)) + 
        theme_minimal() +
        ggtitle("SD by AERONET stations") +
        theme(legend.position ='right')

@@

## adjust for spline trend

```{r, eval=F}
library("splines")
site_selected <- aod_conus[,.N, by = site][N>3330, site]
aod3 <- aod_conus[site == site_selected[1]]
aod3[, dayint:=as.integer(date)]
aod3[, gam(AOD_470nm ~ s(dayint))]
fit0 <- smooth.spline(x = aod3$dayint, y = aod3$AOD_470nm)

plot(aod3[,.(dayint, AOD_470nm)], cex = 0.1)
lines(fit0, col = 'red')

@@

>>>>

---
title: "AODV3 Level2 AOD2.0 from Aeronet Day Variation"
author: "Yang Liu"
date: "7/24/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 8, fig.height = 4)
```

```{r}
source("~/liuyanguu/CWV/Code/cwv_funcs.R")
source(here::here("Code", "funcs_for_plots.R"))
#

@@

compares to figures in https://docs.google.com/document/d/1a1QlET1nDEPf7Sm68iTLgjI3-Quky-Flkib54AaUzpg/edit

@@

@output dataset has dailyN: observations with in a day (at least 20)
@sd_mean and sd_rmse are daily mean and sd

@@

# year > 2000
aod_conus <- aod20_data_pred470[year(date)>=2000][site%in%aeronet_conus_dt$Site_Name][!is.na(AOD_470nm)]
aod_conus[, year := year(date)]
# move the hours 0, 1, 2, attach the earlier date
aod_conus[, hour := as.numeric(hour)]
aod_conus[hour%in%c(0,1,2), `:=`(hour =hour+24, date = date - 1)]

# aod_conus <- aod_conus[year(date)>2000]
aod_conus[,datetime:=date]
aod_conus[,date:=as.Date(date, "%Y-%m-%d")]
aod_conus[,dailyN:=.N, by = .(site, date)] # obs in a day by site,day
aod_conus <- aod_conus[dailyN>=20,]
summary(aod_conus[,.N,by = .(site, date)][,N])
aod_conus[,daysN:= uniqueN(date), by = site] # lasting days by site

@@

@@

* July has the most observations. Ndays are similar, what hours are shorter in winter

```{r}
# do winter have fewer observations per hour? 
aod_conus_hour <- aod_conus[,.(Nobs = .N, Ndays = uniqueN(date)), by = .(season, hour)]  
g1 = ggplot(aod_conus_hour, aes(x = hour, y = Ndays, color = season)) + geom_line() + theme_classic()
g2 = ggplot(aod_conus_hour, aes(x = hour, y = Nobs, color = season)) + geom_line() + theme_classic()
gridExtra::grid.arrange(grobs = list(g1, g2), ncol = 2, top = "By hour: similar days, fewer obs if daytime is short")

@@

## Median Daily SD(AOD)

### calculating using all data

Within year 2000 - 2019 (97% of the total AOD data in CONUS), I calculated sd(AOD_470nm) by site and date. Then take the median.  @mediansd = median(sd_rmse)

```{r,  fig.width=12, fig.height=5}
# summary1: aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[,.(lat, lon, dailyN, daysN, month, season, 
                            mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
                            year = year(date)), by = .(site, date)])
# 290 sites: aod2 [290 * 6]
aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site)])[yearmax>=2018]

# by season

aod3 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site, season)])[yearmax>=2018]

@@

Among 290 stations, 87 have data from 2018 and 2019 (Figure Left).   
@daysN: days of the station with data

* The table shows daysN > 1500 stations.  
* `r aod2[daysN>=500,.N]` with `daysN` > 500 days are plotted  
* `r aod2[daysN>=1000,.N]` with `daysN` > 1000 days and SD(AOD)>0.013(the median)   are labelled in the map. Some other large stations are also labelled in the scatterplot.  
* So in the map, we label the top half of the long-lasting stations with at least 1000 days, and still on at least in 2018.

```{r, fig.width=12, fig.height=5}
# print results with daysN > 500 
setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2[daysN>1500])

plotAODscatter <- function(aod_data, bySeason = F){
  g1  <-  ggplot(data = aod_data[daysN>=500], aes(x = daysN, y = mediansd))+
      geom_point(size = 1, alpha = 1) + 
      theme_classic() + 
      ggrepel::geom_label_repel(data = aod_data[(daysN>1000&mediansd>0.013)|daysN>1500], aes(x = daysN, y = mediansd, label = site)) + 
      ggtitle("Median daily SD(AOD) vs. Total days by Station", subtitle = ("Stations with latest year >=2018")) + 
      labs(x = "Number of Observations", y = "Daily SD(AOD)")
  if(bySeason) {g1 <- g1 + facet_wrap(~season)}
  g1
}

plotAODmap <- function(aod_data, bySeason = F){
  g2 <- ggplot() +
          geom_polygon(data = map_data("state"), 
                       aes(x = long, y = lat, group = group), 
                       fill = 'NA', color = "grey") + 
          geom_point(data = aod_data[daysN>=500],
                     aes(x=lon, y=lat, color = mediansd, size = daysN),
                     alpha = 0.8) + 
          # coord_map("albers", parameters = list(at0 = 45.5, lat1 = 29.5), expand = T) + 
          scale_size_area(max_size = 12) + 
          scale_colour_viridis_c() + 
          labs(x = "Longitude", y = "Latitude", color= "Daily\nSD(AOD)", 
               size = "Number of\nDays") +
          ggrepel::geom_label_repel(data = aod_data[daysN>1000&mediansd>0.013], 
                                    aes(x = lon, y = lat, label = site)) + 
          theme_minimal() +
          ggtitle("SD by AERONET stations") +
          theme(legend.position ='right') 
  if(bySeason) g2 <- g2 + facet_wrap(~season)
  return(g2)
}
p1 <- plotAODscatter(aod2)
p2 <- plotAODmap(aod2)
gridExtra::grid.arrange(grobs = list(p1, p2), ncol = 2)

@@

By season

```{r, fig.width = 12, fig.height = 10}
plotAODscatter(aod3, bySeason = T)
plotAODmap(aod3, bySeason = T)
```

### Calculating using past 3 years: 2016 - 2019

```{r}
# summary2: limiting the years! aod1 calculates the mean and sd (SD) by day 
aod1 <- unique(aod_conus[year>=2016,.(lat, lon, dailyN, daysN, month, season, 
                            mean_aod = mean(AOD_470nm), sd_aod = sd(AOD_470nm),
                            year = year(date)), by = .(site, date)])
# 290 sites: aod2 [290 * 6]
aod2 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site)])[yearmax>=2018]

# by season

aod3 <- unique(aod1[,.(mediansd = median(sd_aod), Nobs = sum(dailyN),
                       yearmax = max(year), daysN, lon, lat),by = .(site, season)])[yearmax>=2018]

@@

```{r, fig.width=12, fig.height=5}
# print results with daysN > 500 
setorder(aod2, -mediansd)
rmarkdown::paged_table(aod2)
p1 <- plotAODscatter(aod2)
p2 <- plotAODmap(aod2)
gridExtra::grid.arrange(grobs = list(p1, p2), ncol = 2)
```

By season

```{r, fig.width = 12, fig.height = 10}
plotAODscatter(aod3, bySeason = T)
plotAODmap(aod3, bySeason = T)
```

## Trend within day

### GSFC, Wallops (high sd) vs. "Sevilleta", "White_Sands_HELSTF" (low sd)

```{r, fig.width = 12, fig.height = 10}
# substract the daily mean -> 
aod_conus[, dailymean := mean(AOD_470nm), by = .(site, date)]
aod_conus[, aod_new := AOD_470nm - dailymean]
aod_conus2 <- aod_conus[site%in%c("Wallops", "GSFC",
                                  "Sevilleta", "White_Sands_HELSTF"),]
 # "MD_Science_Center", "SERC",

# by hour 
ggplot(dilute.data(aod_conus2,10), aes(x = hour , y = aod_new)) + 
  geom_point(size =1, alpha = 0.8) + 
  geom_smooth(se= F, method = "gam", formula = y~s(x, bs = "cs")) + 
  facet_wrap(~site, ncol = 2) +
  theme_classic()

# by date
ggplot(dilute.data(aod_conus2,10), aes(x = date , y = aod_new)) + 
  geom_point(size = 1, alpha = 0.8) + 
  geom_smooth(se= F, method = "gam", formula = y~s(x, bs = "cs")) + 
  facet_wrap(~site, ncol = 2) +
  theme_classic()

@@

```{r }
# show daily trend: AOD vs date, all data
ggplot(data = aod1, aes(x = date, y = sd_aod, color = site)) +
  geom_point(size = 1) + 
  theme_classic() + theme(legend.position = "none") + 
  labs(y = "Daily SD(AOD) by station", x = "Date") + 
  ggtitle("Daily variation by station across time", subtitle = "(# Daily obs >20)")

### 2019-07-30 Code/04_AOD_join_MAIAC.Rmd
check within day variation

---
title: "AOD collocated with MAIAC"
author: "Yang Liu"
date: "07/16/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 7, fig.height = 4.5)
```

@@

@@

Note 7/22, for sure not the best way here, but it can work. 
1. Get geometry for buffers as a list: aod_buffers_list  
2. `st_intersects` buffers to reference grid, get a binary geometry list: join_list
3. get list of dataset of grid within each buffer: ref_list
4. Get geometry for points (site location) as a list: site_list2
5. Within each ref_list (a pie of points), Use `mapply` or a for loop to calculate distance from each grid point to the center (station)

@@

  dist0 <- c(10, 30, 90, 270)
    calculate.vars.in.buffer <- function(distx){
      aod_join_buffer_new <- aod_join_buffer2[dist_km < distx, .(nonmissing = mean(!is.na(AOD_470_buffer)), AOD_buffer_mean = mean(AOD_470_buffer, na.rm = T)),
                              by =.(idLSTpair0, day)]
      setnames(aod_join_buffer_new, c("nonmissing", "AOD_buffer_mean"), 
               paste0(c("pNonNAAOD", "Mean_AOD"),distx, "km"))
      setkey(aod_join_buffer_new, idLSTpair0, day)
      return(aod_join_buffer_new)
    }
  new_varlist <- invisible(lapply(dist0, calculate.vars.in.buffer))
  new_vars <- Reduce(merge, new_varlist)
  setkey(new_vars, idLSTpair0, day)
  setkey(aod_join_MODIS, idLSTpair0, day)
  aod_join_MODIS <- new_vars[aod_join_MODIS]
  return(aod_join_MODIS)
}
aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
head(aod_lst_terra)
# AOD - AOD mean, do later 
saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")

aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
aod_lst_aqua <- aod.MODIS.newVars(aod_lst_aqua, lst_aqua)
head(aod_lst_aqua)
saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")

>>>>

---
title: "AOD Collocated with MAIAC: New Variables"
author: "Yang Liu"
date: "07/16/2019"
output:
  html_document:
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message=FALSE, warning=FALSE,
                      error = F, fig.width = 7, fig.height = 4.5)
```

@@

## Load the collocated dataset

@@

Note 7/22, for sure could be better ways, but it can work. 
1. Get geometry for buffers as a list: aod_buffers_list  
2. `st_intersects` buffers to reference grid, get a binary geometry list: join_list
3. get list of dataset of grid within each buffer: ref_list
4. Get geometry for points (site location) as a list: site_list2
5. Within each ref_list (a pie of points), Use `mapply` or a for loop to calculate distance from each grid point to the center (station)

@@

  dist0 <- c(10, 30, 90, 270)
    calculate.vars.in.buffer <- function(distx){
      aod_join_buffer_new <- aod_join_buffer2[dist_km < distx, .(nonmissing = mean(!is.na(AOD_470_buffer)), AOD_buffer_mean = mean(AOD_470_buffer, na.rm = T)),
                              by =.(idLSTpair0, day)]
      setnames(aod_join_buffer_new, c("nonmissing", "AOD_buffer_mean"), 
               paste0(c("pNonNAAOD", "Mean_AOD"),distx, "km"))
      setkey(aod_join_buffer_new, idLSTpair0, day)
      return(aod_join_buffer_new)
    }
  new_varlist <- invisible(lapply(dist0, calculate.vars.in.buffer))
  new_vars <- Reduce(merge, new_varlist)
  setkey(new_vars, idLSTpair0, day)
  setkey(aod_join_MODIS, idLSTpair0, day)
  aod_join_MODIS <- new_vars[aod_join_MODIS]
  return(aod_join_MODIS)
}

## Terra
aod_lst_terra <- aod.MODIS.newVars(aod_lst_terra, lst_terra)
head(aod_lst_terra)
# AOD - AOD mean, do later 
saveRDS(aod_lst_terra, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_terra_newVars.rds")

## Aqua
aod_lst_aqua <- readRDS("~/Intermediate/AOD_data/190629_LSTjoinMCD19_t_a.rds")$aod_lst_aqua
aod_lst_aqua <- aod.MODIS.newVars(aod_lst_aqua, lst_aqua)
head(aod_lst_aqua)
saveRDS(aod_lst_aqua, "~/Intermediate/AOD_data/190724_LSTjoinMCD19_aqua_newVars.rds")
